Bluetooth Loader for mruby Bytecode \\ in Multiple Virtual Machine Environment

### 2016-05-25 CPSNA2016/paper.tex
add CPSNA

%
% File: thesis.tex
% Created: 月 11 30 08:00 午後 2015 東京 (標準時)
% Last Change: 月 11 30 08:00 午後 2015 東京 (標準時)
%
\input{dummy/dummy.tex}
\documentclass{sig-alternate-05-2015}
%\usepackage[dvipdfmx]{graphicx}
\usepackage{amssymb}
\usepackage{enumerate,cite,url}
\usepackage{listings,jlisting}
\lstset{%
    language={c},%
    basicstyle={\small},%
    identifierstyle={\small},%
    commentstyle={\small\itshape},%
    keywordstyle={\small},%\bfseries},%
    ndkeywordstyle={\small},%
    stringstyle={\small\it},
    frame={tb},
    breaklines=true,
    columns=[l]{fullflexible},%
    numbers=left,%
    xrightmargin=0zw,%
    xleftmargin=3zw,%
    numberstyle={\scriptsize},%
    stepnumber=1,
    numbersep=1zw,%
    %lineskip=-0.5ex%
}

% Copyright
\setcopyright{acmcopyright}
% DOI
\doi{10.475/123_4}
% ISBN
\isbn{123-4567-24-567/08/06}
%Conference
\conferenceinfo{EMSOFT '16}{October 2--7, 2016, Pittsburgh, USA}
\acmPrice{\$15.00}

\title{Bluetooth Loader for mruby Bytecode \\ in Multiple Virtual Machine Environment}
\numberofauthors{3}
% \author{
% \alignauthor
% Takuro Yamamoto\\
%     \affaddr{Graduate School of Engineering Science, Osaka University}\\
%     \affaddr{Osaka, JAPAN}\\
%        % \affaddr{Wallamaloo, New Zealand}\\
%        % \email{trovato@corporation.com}
% \alignauthor
% Hiroshi Oyama\\
%     \affaddr{OKUMA Corporation}\\
%     \affaddr{Aichi, JAPAN}\\
%        % \affaddr{Dublin, Ohio 43017-6221}\\
%        % \email{webmaster@marysville-ohio.com}
% \alignauthor
% Takuya Azumi\\
%     \affaddr{Graduate School of Engineering Science, Osaka University}\\
%     \affaddr{Osaka, JAPAN}\\
%        % \affaddr{Hekla, Iceland}\\
%        % \email{larst@affiliation.org}
% }
%\author{Takuro Yamamoto}
%\date{\today}

@@

\keywords{embedded software; scripting language; component-based development}
\section{Introduction}
Currently, embedded systems, e.g., Internet of Things applications, must demonstrate high quality and high performance.
This requirement has led to an increase in their complexity and scale; moreover, these systems need to have low production costs and short development cycles.

Complex and large-scale software systems can be developed efficiently by using component-based techniques \cite{par:Crnkovic}, \cite{par:CBD}.
Component-Based Development (CBD) is a design technique that can be applied to reusable software development.
% Complex and large scale software systems can be developed efficiently using component-based techniques.
% That is because software componentization provides high-reusability.
Verification of component-based systems has been extensively researched \cite{par:Blaming}, \cite{par:Verification}.
Individual component diagrams enable the visualization of an entire system.
%It also visualize the entire system due to component diagrams, and makes a system flexible in extensions and specification changes.
In addition, component-based systems are flexible with regard to extensibility and specification changes.
The TOPPERS embedded component system (TECS) \cite{par:TECS}, AUTOSAR \cite{url:AUTOSAR}, and SaveCCM \cite{par:SAVEapproach} are typical CBD tools for embedded systems.

@@

{\mybf Contributions}: The proposed framework provides the following contributions:
\begin{enumerate}
\item {\mybf Improved software development efficiency.}
Developers do not need to rewrite a storage/ROM device and restart an RTOS.
The Bluetooth loader supports continuous loading, which reduces Bluetooth set-up time (i.e., pairing).
\item {\mybf Execution of multiple mruby programs concurrently or in parallel.}
Developers can implement multiple tasks without RTOS knowledge because the RiteVM scheduler switches tasks cyclically.
\item {\mybf Synchronized execution of multiple RiteVM tasks.}
The proposed framework synchronizes multiple RiteVM tasks (i.e., mruby applications).
\item {\mybf Benefits of CBD:}
The paper focuses on the benefits of CBD and provides specific examples.
\end{enumerate}

{\mybf Organization}: The reminder of this paper is organized as follows.
Section \ref{sec:Background} introduces the basic technologies, i.e., mruby, TECS, and mruby on TECS.
Section \ref{sec:Design and Implementation} describes the design and implementation of the proposed framework.
Section \ref{sec:Evaluation} evaluates the proposed framework.
Related work is discussed in Section \ref{sec:Related work}.
Conclusions and suggestions for future work are presented in Section \ref{sec:Conclusion}.

\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/proposed.eps}
    \vspace{0.5mm}
    \caption{System model of the proposed framework}
    \vspace{1mm}
    \label{fig:proposed}
\end{figure}

@@

mruby, which retains the usability and readability of Ruby, requires fewer resources, and thus, is suitable for embedded systems.
In addition, mruby includes a VM mechanism, and thus, mruby programs can run on any operating system as long as a VM is implemented.
The mruby/RiteVM mechanism is shown in Figure \ref{fig:mruby}.
The mruby compiler translates an mruby code into a bytecode, which can be interpreted by a RiteVM; thus, mruby programs can be executed on any target device with a RiteVM.
\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/mruby.eps}
    \vspace{0.5mm}
    \caption{mruby/RiteVM mechanism}
    \vspace{1mm}
    \label{fig:mruby}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/component_diagram.eps}
    % \vspace{0.5mm}
    \caption{Component Diagram}
    \vspace{1mm}
    \label{fig:component}
\end{figure}

\subsubsection{Component Model}
Figure \ref{fig:component} shows a component diagram.
A {\myit cell}, which is an instance of a component in TECS, consists of {\myit entry} ports, {\myit call} ports, attributes and internal variables.
An {\myit entry} port is an interface that provides functions to other {\myit cell}s, and a {\myit call} port is an interface that enables the use of other {\myit cell}'s functions.
A {\myit cell} has one or more {\myit entry} ports and {\myit call} ports.
{\myit Cell} functions are implemented in C.

The type of {\myit entry}/{\myit call} port is defined by a {\myit signature}, which is a set of functions.
A {\myit signature} is the interface definition of a {\myit cell}.
The {\myit cell}'s  {\myit call} port can be connected to the {\myit entry} port of another {\myit cell} by the same {\myit signature}.
Here, {\myit celltype} defines one or more {\myit call}/{\myit entry} ports, attributes, and internal variables of a {\myit cell}.

\subsubsection{Component Description}
In TECS, components are described by {\myit signature}, {\myit celltype}, and build written in component description language (CDL).
% TECS code is written in CDL (component description language) file.
These components are described as follows.

\begin{description}
    \item[{\mybf Signature Description}]\mbox{}\\
        The {\myit signature} defines a {\myit cell} interface.
        The {\myit signature} name follows the keyword {\myit signature} and takes the prefix ``s'' e.g., sMotor (Figure \ref{signature}).
        In TECS, to clarify the function of an interface, specifiers such as [in] and [out] are used, which represent input and output, respectively.
\begin{figure}[t]
\centering
\begin{lstlisting}
signature sMotor {
    int32_t getCounts( void );
    ER resetCounts( void );
    ER setPower( [in]int power );
    ER stop( [in] bool_t brake );
    ER rotate( [in] int degrees, [in] uint32_t speed_abs,
              [in] bool_t blocking );
    void initializePort( [in]int32_t type );
};
\end{lstlisting}
\vspace{1mm}
\caption{Signature Description}
\vspace{1mm}
\label{signature}
\end{figure}
    \item[{\mybf Celltype Description}]\mbox{}\\
        The {\myit celltype} defines {\myit entry} ports, {\myit call} ports, attributes, and variables.
        A {\myit celltype} name with the prefix ``t'' follows the keyword {\myit celltype}, e.g., tCaller (Figure \ref{celltype}).
        To define {\myit entry} ports, a {\myit signature}, e.g., sMotor, and an {\myit entry} port name, e.g., eMotor, follow the keyword {\myit entry}.
        {\myit Call} ports are defined similarly.
        Attributes and variables follow the keywords {\myit attr} and {\myit var}, respectively.
\begin{figure}[t]
\centering
\begin{lstlisting}
celltype tCaller {
    call sMotor cMotor;
};
celltype tMotor {
    entry sMotor eMotor;
    attr {
        int32_t port;
    };
    var {
        int32_t currentSpeed = 0;
    };
};
\end{lstlisting}
\vspace{1mm}
\caption{Celltype Description}
\vspace{1mm}
\label{celltype}
\end{figure}
    \item[{\mybf Build Description}]\mbox{}\\
        The build description is used to instantiate and connect {\myit cell}s.
        Figure \ref{build} shows an example of a build description.
        A {\myit celltype} name and {\myit cell} name, e.g., tMotor and Motor, respectively, follow the keyword {\myit cell}.
        To compose {\myit cell}s, a {\myit call} port, {\myit cell}'s name, and an {\myit entry} port are described in that order.
        In Figure \ref{build}, {\myit entry} port eMotor in {\myit cell} Motor is connected to {\myit call} port cMotor in {\myit cell} Caller.
        {\myit C\_EXP} calls macros defined in C files.

\begin{figure}[t]
\centering
\begin{lstlisting}
cell tMotor Motor {
    port = C_EXP("PORT_A");
};
cell tCaller Caller {
    cMotor = Motor.eMotor;
};
\end{lstlisting}
\vspace{1mm}
\caption{Build Description}
\vspace{1mm}
\label{build}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/mrubyontecs.eps}
    \vspace{0.5mm}
\caption{System model of existing mruby on TECS}
    \vspace{1mm}
\label{fig:mrubyontecs}
\end{figure}

@@

Figure \ref{fig:mruby_TECS_bridge} illustrates an mruby-TECS bridge used to control a motor.
The left side of BridgeMotor belongs to the mruby program.
The right side of BridgeMotor belongs to TECS component.
\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/mruby_TECS_bridge.eps}
    \vspace{0.5mm}
\caption{mruby-TECS bridge}
    \vspace{1mm}
\label{fig:mruby_TECS_bridge}
\end{figure}

The mruby-TECS bridge generates a {\myit celltype}, which is called from the mruby code, and an mruby class, which corresponds to a developer-specified TECS component to invoke a C function from the mruby program.

The generated mruby-TECS bridge supports registration of classes and methods for mruby.
Methods in an mruby class are defined by generation codes for an mruby-TECS bridge, such as setPower and stop.
Thus, when a method is called in an mruby program, the mruby-TECS bridge calls the function defined in the TECS component such as a Motor {\myit cell}.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/system_model.eps}
    \vspace{0.5mm}
\caption{Detailed system model of the proposed framework}
    \vspace{1mm}
\label{fig:system_model}
\end{figure}

@@

\subsubsection{RiteVM Component with Bluetooth Loader for mruby Bytecode}
\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/component_bluetooth.eps}
    \vspace{0.5mm}
\caption{Component diagram of Bluetooth loader for mruby bytecode}
    \vspace{1mm}
\label{fig:component_bluetooth}
\end{figure}
The proposed framework provides a RiteVM with a Bluetooth loader for mruby bytecode as a TECS component.
This component is an extension of the RiteVM component described in \cite{par:mrubyonTECS}.
It receives bytecodes via Bluetooth and manages the RiteVM configuration, such as generating mruby library bytecodes automatically.
This generated bytecode is prepared beforehand on the storage/ROM device and differs from the bytecode transferred with Bluetooth.

Figure \ref{fig:component_bluetooth} shows a component diagram of MrubyTask1 and RiteVMBluetooth1 {\myit cell}s.
The MrubyTask1 {\myit cell} is a componentized task of the RTOS (TOPPERS/HRP2 \cite{url:HRP2}, \cite{par:hr-tecs}).
The RiteVMBluetooth1 {\myit cell} is the RiteVM component with the Bluetooth loader for mruby bytecode.
Bytecode on the host is transferred and received.
In this framework, ZMODEM \cite{par:zmodem} is used as the binary transfer protocol.

\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/control_flow.eps}
    \vspace{0.5mm}
\caption{Process flow of Bluetooth loader for mruby bytecode}
    \vspace{1mm}
\label{fig:control_flow}
\end{figure}

\begin{figure}[t]
\centering
\begin{lstlisting}
/* tRiteVMBluetooth.cdl */
celltype tRiteVMBluetooth1 {
    entry sTaskBody eMrubyBody;
    [optional] call sEventflag cEventflag[];
    [optional] call sSemaphore cSemaphore;
    attr {
      [omit]char_t *mrubyLib;
      char_t *irepLib = 
                    C_EXP("&$cell_global$_irep");
      uint32_t irepAppSize = 
                    C_EXP( BUFFER_SIZE );
      FLGPTN setptn;
    };
    var {
        mrb_state *mrb;
        mrbc_context *context;
        [size_is(irepAppSize)] uint8_t *irepApp;
    };
};
\end{lstlisting}
\vspace{1mm}
\caption{Celltype description for RiteVM with Bluetooth loader for mruby bytecode}
\vspace{1mm}
\label{celltype_mrubybluetooth}
\end{figure}

@@

\end{lstlisting}
\vspace{1mm}
\caption{Main code for RiteVM with Bluetooth loader for mruby bytecode}
\vspace{1mm}
\label{maincode_mrubybluetooth}
\end{figure}

@@

First, the Bluetooth loader receives the mruby application bytecode from the host (Figure \ref{fig:control_flow}(A); lines 6-7 in Figure \ref{maincode_mrubybluetooth}).
The bytecode is stored in a component variable, such as {\myit VAR\_irepApp}, as shown in Figure \ref{celltype_mrubybluetooth}.
This process is exclusively carried out by the semaphore.

Second, {\myit mrb\_state} and {\myit mrbc\_context} pointers, and mruby-TECS bridges are initialized (Figure \ref{fig:control_flow}(B); lines 9-14 in Figure \ref{maincode_mrubybluetooth}).
{\myit VAR\_mrb} and {\myit VAR\_context} show the variables of the {\myit cell}.
{\myit mrb\_state} is a set of states and global variables used in mruby.
The synchronization of multiple tasks is performed in this processing phase.
The RiteVM that finishes execution at this point waits for the another RiteVM to finish loading and initialization.

Third, the RiteVM reads the bytecode of mruby libraries (Figure \ref{fig:control_flow}(C); lines 17-19 in Figure \ref{maincode_mrubybluetooth}).
The mruby libraries are a set of Ruby classes, such as the motor and sensor classes.
For example, the motor class defines methods to rotate and stop a motor.
The tRiteVMBluetooth {\myit cell} has the attributes shown in Figure \ref{celltype_mrubybluetooth}.
{\myit ATTR} indicates an attribute which is a fixed value that cannot be rewritten, unlike {\myit VAR}.
The {\myit mrubyLib} indicates the program files of the mruby libraries, and is an attribute because mruby libraries are not modified in the proposed development process.
Here, {\myit [omit]} is only used for the TECS generator; thus, the attribute {\myit mrubyLib} does not consume memory.
{\myit irepLib} is the pointer of the array in which the bytecode of mruby libraries is stored.
To summarize, the bytecode of mruby libraries is stored as an attribute of the component during the first compilation.

Fourth, the RiteVM reads the bytecode of the mruby application transferred via Bluetooth (Figure \ref{fig:control_flow}(D); lines 20-22 in Figure \ref{maincode_mrubybluetooth}).
The mruby application bytecode is stored in an array of {\myit irepApp}, which differs from the array that holds the mruby library bytecode.
Note that two bytecodes are read separately in the RiteVM.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/rotateReadyQueue.eps}
    \vspace{0.5mm}
\caption{RiteVM scheduler design}
    \vspace{1mm}
\label{fig:rotateReadyQueue}
\end{figure}

\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/cyclic_handler.eps}
    % \vspace{0.5mm}
\caption{Component diagram of RiteVM scheduler}
    \vspace{1mm}
\label{fig:cyclic_handler}
\end{figure}

@@

One approach for multitasking is a co-routine, which is a cooperative thread scheduled by developers with functions such as {\myit resume} and {\myit yield} (the Ruby co-routine is defined in the Fiber class \cite{url:co-routine}).
A co-routine is non-preemptive multitasking, which does not receive OS support because developers must switch tasks manually.
A co-routine cannot take advantage of multicore processing.

As another method, {\myit delay()}, which is a service call of $\mu$ITRON, can be used for multitasking.
This service call delays the execution of its own task for the time of the argument.
{\myit delay()} is needed when scheduling fixed-priority tasks.
However, the programming applied to {\myit delay()} is difficult to use with fair scheduling.

@@

\subsubsection{RiteVM Scheduler Design}
A RiteVM scheduler is a periodic handler, and {\myit rotateReadyQueue}, which is a service call of $\mu$ITRON to switch tasks with equal priority, is implemented as the main process of the handler.
In other words, the RiteVM scheduler calls {\myit rotateReadyQueue} cyclically.
The design of the RiteVM scheduler is shown in Figure \ref{fig:rotateReadyQueue}. 
{\myit rotateReadyQueue} is described as follows.

@@

When {\myit rotateReadyQueue} is called, tasks with equal priority are switched as shown in Figure \ref{fig:rotateReadyQueue}.
Note that the argument of {\myit rotateReadyQueue} is the priority.

In addition, {\myit rotateReadyQueue} can be performed if the number of tasks is more than two.
For example, three tasks are in the order task1, task2, and task3.
In this case, the order is rotated to task2, task3, and task1 when {\myit rotateReadyQueue} is called.

@@

\subsection{Utilization of Component-Based Development}
% This section describes the design applying component-based development.
In the proposed framework, RiteVMs, the RiteVM scheduler, and Eventflags are implemented as components.
Therefore, developers can add, remove, or reuse these components easily.
For example, if the RiteVM scheduler is not necessary for the software, developers should comment out only the CDL file, e.g., {\myit //import(<tRiteVMScheduler.cdl>);}.
CBD eliminates the need for developers to rewrite kernel configuration files.

>>>>

%
% File: thesis.tex
% Created: 月 11 30 08:00 午後 2015 東京 (標準時)
% Last Change: 月 11 30 08:00 午後 2015 東京 (標準時)
%
\input{dummy/dummy.tex}
\documentclass[conference]{IEEEtran}
\usepackage[dvipdfmx]{graphicx}
\usepackage{amssymb}
\usepackage{enumerate,cite,url}
\usepackage{listings,jlisting}
\lstset{%
    language={c},%
    basicstyle={\small},%
    identifierstyle={\small},%
    commentstyle={\small\itshape},%
    keywordstyle={\small},%\bfseries},%
    ndkeywordstyle={\small},%
    stringstyle={\small\it},
    frame={tb},
    breaklines=true,
    columns=[l]{fullflexible},%
    numbers=left,%
    xrightmargin=0zw,%
    xleftmargin=3zw,%
    numberstyle={\scriptsize},%
    stepnumber=1,
    numbersep=1zw,%
    %lineskip=-0.5ex%
}

\title{Bluetooth Loader for mruby Bytecode \\ in Multiple Virtual Machine Environment}
\author{\IEEEauthorblockN{Takuro Yamamoto}\IEEEauthorblockA{Graduate School of Engineering Science,\\Osaka University}
\and\IEEEauthorblockN{Hiroshi Oyama}\IEEEauthorblockA{OKUMA Corporation}
\and\IEEEauthorblockN{Takuya Azumi}\IEEEauthorblockA{Graduate School of Engineering Science,\\Osaka University}
}

@@

\section{Introduction}
Currently, embedded systems, e.g., Internet of Things applications, must demonstrate high quality and high performance.
This requirement has led to an increase in their complexity and scale; moreover, these systems need to have low production costs and short development cycles.

Complex and large-scale software systems can be developed efficiently by using component-based techniques \cite{par:Crnkovic}, \cite{par:CBD}.
Component-Based Development (CBD) is a design technique that can be applied to reusable software development.
% Complex and large scale software systems can be developed efficiently using component-based techniques.
% That is because software componentization provides high-reusability.
% Verification of component-based systems has been extensively researched \cite{par:Blaming}, \cite{par:Verification}.
Individual component diagrams enable the visualization of an entire system.
%It also visualize the entire system due to component diagrams, and makes a system flexible in extensions and specification changes.
In addition, component-based systems are flexible with regard to extensibility and specification changes.
The TOPPERS embedded component system (TECS) \cite{par:TECS}, AUTOSAR \cite{url:AUTOSAR}, and SaveCCM \cite{par:SAVEapproach} are typical CBD tools for embedded systems.

@@

{\bf Contributions}: The proposed framework provides the following contributions:
\begin{enumerate}
\item {\bf Improved software development efficiency.}
Developers do not need to rewrite a storage/ROM device and restart an RTOS.
The Bluetooth loader supports continuous loading, which reduces Bluetooth set-up time (i.e., pairing).
\item {\bf Execution of multiple mruby programs concurrently or in parallel.}
Developers can implement multiple tasks without RTOS knowledge because the RiteVM scheduler switches tasks cyclically.
\item {\bf Synchronized execution of multiple RiteVM tasks.}
The proposed framework synchronizes multiple RiteVM tasks (i.e., mruby applications).
\item {\bf Benefits of CBD:}
The paper focuses on the benefits of CBD and provides specific examples.
\end{enumerate}

{\bf Organization}: The reminder of this paper is organized as follows.
Section \ref{sec:Background} introduces the basic technologies, i.e., mruby, TECS, and mruby on TECS.
Section \ref{sec:Design and Implementation} describes the design and implementation of the proposed framework.
Section \ref{sec:Evaluation} evaluates the proposed framework.
Related work is discussed in Section \ref{sec:Related work}.
Conclusions and suggestions for future work are presented in Section \ref{sec:Conclusion}.

\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/proposed.eps}
    \caption{System model of the proposed framework}
    \label{fig:proposed}
\end{figure}

@@

mruby, which retains the usability and readability of Ruby, requires fewer resources, and thus, is suitable for embedded systems.
In addition, mruby includes a VM mechanism, and thus, mruby programs can run on any operating system as long as a VM is implemented.
The mruby/RiteVM mechanism is shown in Figure \ref{fig:mruby}.
The mruby compiler translates an mruby code into a bytecode, which can be interpreted by a RiteVM; thus, mruby programs can be executed on any target device with a RiteVM.
\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/mruby.eps}
    \caption{mruby/RiteVM mechanism}
    \label{fig:mruby}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/component_diagram.eps}
    \caption{Component Diagram}
    \label{fig:component}
\end{figure}

\subsubsection{Component Model}
Figure \ref{fig:component} shows a component diagram.
A {\it cell}, which is an instance of a component in TECS, consists of {\it entry} ports, {\it call} ports, attributes and internal variables.
An {\it entry} port is an interface that provides functions to other {\it cell}s, and a {\it call} port is an interface that enables the use of other {\it cell}'s functions.
A {\it cell} has one or more {\it entry} ports and {\it call} ports.
{\it Cell} functions are implemented in C.

The type of {\it entry}/{\it call} port is defined by a {\it signature}, which is a set of functions.
A {\it signature} is the interface definition of a {\it cell}.
The {\it cell}'s  {\it call} port can be connected to the {\it entry} port of another {\it cell} by the same {\it signature}.
Here, {\it celltype} defines one or more {\it call}/{\it entry} ports, attributes, and internal variables of a {\it cell}.

\subsubsection{Component Description}
In TECS, components are described by {\it signature}, {\it celltype}, and build written in component description language (CDL).
% TECS code is written in CDL (component description language) file.
These components are described as follows.

\begin{description}
    \item[{\bf Signature Description}]\mbox{}\\
        The {\it signature} defines a {\it cell} interface.
        The {\it signature} name follows the keyword {\it signature} and takes the prefix ``s'' e.g., sMotor (Figure \ref{signature}).
        In TECS, to clarify the function of an interface, specifiers such as [in] and [out] are used, which represent input and output, respectively.
\begin{figure}[t]
\centering
\begin{lstlisting}
signature sMotor {
    int32_t getCounts( void );
    ER resetCounts( void );
    ER setPower( [in]int power );
    ER stop( [in] bool_t brake );
    ER rotate( [in] int degrees, [in] uint32_t speed_abs,
              [in] bool_t blocking );
    void initializePort( [in]int32_t type );
};
\end{lstlisting}
\caption{Signature Description}
\label{signature}
\end{figure}
    \item[{\bf Celltype Description}]\mbox{}\\
        The {\it celltype} defines {\it entry} ports, {\it call} ports, attributes, and variables.
        A {\it celltype} name with the prefix ``t'' follows the keyword {\it celltype}, e.g., tCaller (Figure \ref{celltype}).
        To define {\it entry} ports, a {\it signature}, e.g., sMotor, and an {\it entry} port name, e.g., eMotor, follow the keyword {\it entry}.
        {\it Call} ports are defined similarly.
        Attributes and variables follow the keywords {\it attr} and {\it var}, respectively.
\begin{figure}[t]
\centering
\begin{lstlisting}
celltype tCaller {
    call sMotor cMotor;
};
celltype tMotor {
    entry sMotor eMotor;
    attr {
        int32_t port;
    };
    var {
        int32_t currentSpeed = 0;
    };
};
\end{lstlisting}
\caption{Celltype Description}
\label{celltype}
\end{figure}
    \item[{\bf Build Description}]\mbox{}\\
        The build description is used to instantiate and connect {\it cell}s.
        Figure \ref{build} shows an example of a build description.
        A {\it celltype} name and {\it cell} name, e.g., tMotor and Motor, respectively, follow the keyword {\it cell}.
        To compose {\it cell}s, a {\it call} port, {\it cell}'s name, and an {\it entry} port are described in that order.
        In Figure \ref{build}, {\it entry} port eMotor in {\it cell} Motor is connected to {\it call} port cMotor in {\it cell} Caller.
        {\it C\_EXP} calls macros defined in C files.

\begin{figure}[t]
\centering
\begin{lstlisting}
cell tMotor Motor {
    port = C_EXP("PORT_A");
};
cell tCaller Caller {
    cMotor = Motor.eMotor;
};
\end{lstlisting}
\caption{Build Description}
\label{build}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/mrubyontecs.eps}
\caption{System model of existing mruby on TECS}
\label{fig:mrubyontecs}
\end{figure}

@@

Figure \ref{fig:mruby_TECS_bridge} illustrates an mruby-TECS bridge used to control a motor.
The left side of BridgeMotor belongs to the mruby program.
The right side of BridgeMotor belongs to TECS component.
\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/mruby_TECS_bridge.eps}
\caption{mruby-TECS bridge}
\label{fig:mruby_TECS_bridge}
\end{figure}

The mruby-TECS bridge generates a {\it celltype}, which is called from the mruby code, and an mruby class, which corresponds to a developer-specified TECS component to invoke a C function from the mruby program.

The generated mruby-TECS bridge supports registration of classes and methods for mruby.
Methods in an mruby class are defined by generation codes for an mruby-TECS bridge, such as setPower and stop.
Thus, when a method is called in an mruby program, the mruby-TECS bridge calls the function defined in the TECS component such as a Motor {\it cell}.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/system_model.eps}
\caption{Detailed system model of the proposed framework}
\label{fig:system_model}
\end{figure}

@@

\subsubsection{RiteVM Component with Bluetooth Loader for mruby Bytecode}
\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/component_bluetooth.eps}
\caption{Component diagram of Bluetooth loader for mruby bytecode}
\label{fig:component_bluetooth}
\end{figure}
The proposed framework provides a RiteVM with a Bluetooth loader for mruby bytecode as a TECS component.
This component is an extension of the RiteVM component described in \cite{par:mrubyonTECS}.
It receives bytecodes via Bluetooth and manages the RiteVM configuration, such as generating mruby library bytecodes automatically.
This generated bytecode is prepared beforehand on the storage/ROM device and differs from the bytecode transferred with Bluetooth.

Figure \ref{fig:component_bluetooth} shows a component diagram of MrubyTask1 and RiteVMBluetooth1 {\it cell}s.
The MrubyTask1 {\it cell} is a componentized task of the RTOS (TOPPERS/HRP2 \cite{url:HRP2}, \cite{par:hr-tecs}).
The RiteVMBluetooth1 {\it cell} is the RiteVM component with the Bluetooth loader for mruby bytecode.
Bytecode on the host is transferred and received.
In this framework, ZMODEM \cite{par:zmodem} is used as the binary transfer protocol.

\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/control_flow.eps}
\caption{Process flow of Bluetooth loader for mruby bytecode}
\label{fig:control_flow}
\end{figure}

\begin{figure}[t]
\centering
\begin{lstlisting}
/* tRiteVMBluetooth.cdl */
celltype tRiteVMBluetooth1 {
    entry sTaskBody eMrubyBody;
    [optional] call sEventflag cEventflag[];
    [optional] call sSemaphore cSemaphore;
    attr {
      [omit]char_t *mrubyLib;
      char_t *irepLib = 
                    C_EXP("&$cell_global$_irep");
      uint32_t irepAppSize = 
                    C_EXP( BUFFER_SIZE );
      FLGPTN setptn;
    };
    var {
        mrb_state *mrb;
        mrbc_context *context;
        [size_is(irepAppSize)] uint8_t *irepApp;
    };
};
\end{lstlisting}
\caption{Celltype description for RiteVM with Bluetooth loader for mruby bytecode}
\label{celltype_mrubybluetooth}
\end{figure}

@@

\end{lstlisting}
\caption{Main code for RiteVM with Bluetooth loader for mruby bytecode}
\label{maincode_mrubybluetooth}
\end{figure}

@@

First, the Bluetooth loader receives the mruby application bytecode from the host (Figure \ref{fig:control_flow}(A); lines 6-7 in Figure \ref{maincode_mrubybluetooth}).
The bytecode is stored in a component variable, such as {\it VAR\_irepApp}, as shown in Figure \ref{celltype_mrubybluetooth}.
This process is exclusively carried out by the semaphore.

Second, {\it mrb\_state} and {\it mrbc\_context} pointers, and mruby-TECS bridges are initialized (Figure \ref{fig:control_flow}(B); lines 9-14 in Figure \ref{maincode_mrubybluetooth}).
{\it VAR\_mrb} and {\it VAR\_context} show the variables of the {\it cell}.
{\it mrb\_state} is a set of states and global variables used in mruby.
The synchronization of multiple tasks is performed in this processing phase.
The RiteVM that finishes execution at this point waits for the another RiteVM to finish loading and initialization.

Third, the RiteVM reads the bytecode of mruby libraries (Figure \ref{fig:control_flow}(C); lines 17-19 in Figure \ref{maincode_mrubybluetooth}).
The mruby libraries are a set of Ruby classes, such as the motor and sensor classes.
For example, the motor class defines methods to rotate and stop a motor.
The tRiteVMBluetooth {\it cell} has the attributes shown in Figure \ref{celltype_mrubybluetooth}.
{\it ATTR} indicates an attribute which is a fixed value that cannot be rewritten, unlike {\it VAR}.
The {\it mrubyLib} indicates the program files of the mruby libraries, and is an attribute because mruby libraries are not modified in the proposed development process.
Here, {\it [omit]} is only used for the TECS generator; thus, the attribute {\it mrubyLib} does not consume memory.
{\it irepLib} is the pointer of the array in which the bytecode of mruby libraries is stored.
To summarize, the bytecode of mruby libraries is stored as an attribute of the component during the first compilation.

Fourth, the RiteVM reads the bytecode of the mruby application transferred via Bluetooth (Figure \ref{fig:control_flow}(D); lines 20-22 in Figure \ref{maincode_mrubybluetooth}).
The mruby application bytecode is stored in an array of {\it irepApp}, which differs from the array that holds the mruby library bytecode.
Note that two bytecodes are read separately in the RiteVM.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/rotateReadyQueue.eps}
\caption{RiteVM scheduler design}
\label{fig:rotateReadyQueue}
\end{figure}

\begin{figure}[t]
    \centering
    \includegraphics[width=8.6cm,clip]{figure/cyclic_handler.eps}
\caption{Component diagram of RiteVM scheduler}
\label{fig:cyclic_handler}
\end{figure}

@@

One approach for multitasking is a co-routine, which is a cooperative thread scheduled by developers with functions such as {\it resume} and {\it yield} (the Ruby co-routine is defined in the Fiber class \cite{url:co-routine}).
A co-routine is non-preemptive multitasking, which does not receive OS support because developers must switch tasks manually.
A co-routine cannot take advantage of multicore processing.

As another method, {\it delay()}, which is a service call of $\mu$ITRON, can be used for multitasking.
This service call delays the execution of its own task for the time of the argument.
{\it delay()} is needed when scheduling fixed-priority tasks.
However, the programming applied to {\it delay()} is difficult to use with fair scheduling.

@@

\subsubsection{RiteVM Scheduler Design}
A RiteVM scheduler is a periodic handler, and {\it rotateReadyQueue}, which is a service call of $\mu$ITRON to switch tasks with equal priority, is implemented as the main process of the handler.
In other words, the RiteVM scheduler calls {\it rotateReadyQueue} cyclically.
The design of the RiteVM scheduler is shown in Figure \ref{fig:rotateReadyQueue}. 
{\it rotateReadyQueue} is described as follows.

@@

When {\it rotateReadyQueue} is called, tasks with equal priority are switched as shown in Figure \ref{fig:rotateReadyQueue}.
Note that the argument of {\it rotateReadyQueue} is the priority.

In addition, {\it rotateReadyQueue} can be performed if the number of tasks is more than two.
For example, three tasks are in the order task1, task2, and task3.
In this case, the order is rotated to task2, task3, and task1 when {\it rotateReadyQueue} is called.

@@

\subsection{Utilization of Component-Based Development}
% This section describes the design applying component-based development.
In the proposed framework, RiteVMs, the RiteVM scheduler, and Eventflags are implemented as components.
Therefore, developers can add, remove, or reuse these components easily.
For example, if the RiteVM scheduler is not necessary for the software, developers should comment out only the CDL file, e.g., {\it //import(<tRiteVMScheduler.cdl>);}.
CBD eliminates the need for developers to rewrite kernel configuration files.

In addition, the code size can be reduced by using CBD. 
In the proposed framework, this advantage is applied in the Eventflag component.
The set pattern and wait pattern are defined as attributes of the component as shown in Figure \ref{fig:Eventflag} (B).
This design, e.g., {\it cEventflag\_set(ATTR\_setPattern)}, enables the program without ``if'' statements and reuses an identical C file.
Developers do not need to modify the C file because the CDL files are prepared according to the number of RiteVMs.
In addition, the Eventflag components are built with {\it [optional]} in TECS.
Here, {\it [optional]} means that the code is run only when the call port is connected.
Note that the C file does not need to be rewritten even if the Eventflag is not used.
